---
title: "NBA salary analysis"
author: "Jackson Yang"
date: '2022-07-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# 1. Load library
```{r}
library(data.table)
library(ggthemes)
library(tidyverse)
library(GGally)
library(plotly)
library(plotly)
library(mde)
library(corrplot)
library(viridis)
library(caTools)
theme_set(theme_few())
```

# 2. Load Data
```{r}
salary <- as_tibble(fread('NBA_season1718_salary.csv'))
ss <- as_tibble(fread('Seasons_Stats.csv'))
```

# 3. Data Overview 
## 3.1 salary
```{r}
head(salary)
summary(salary)
```
**After viewing the head and summary, we have noticed several things**
a).colname "season17_18" should be changed to "salary"; 
b).there is no null values in the salary table; 
c).this table is already ranked based off salary.

## 3.2 season statistics
```{r}
head(ss)
# summary(ss)
```

**After viewing the head and summary, we have noticed several things**
a).this table contains lots of columns as well as NA values.
b). stat data ranges from 1950 to ....
c). there were players who played for different teams within a season
d). completely empty columns: blani/blank2 which should be dropped

# 4. Data Cleaning
## 4.1 Data merge and select
* we want to just use the most recent data 2016-2017 season
* we will merge salary and stats17
* we will get rid of empty columns and keep only columns that 'I' understand :) as a non basketball analytic expert
```{r}
ss17 <- ss %>% filter(Year==2017)
df <- merge(ss17, salary, by.x='Player',by.y = 'Player')
df <- df %>% select(
   Player, 
   Age,
   Team = Tm.y,
   ORB,DRB,AST,STL,BLK,TOV,PTS,
   G,GS,MP,Eff= PER,"TS%",FG,FGA,'FG%',"3P","3P%","2P","2PA","2P%",
   Salary=season17_18)
```
## 4.2 Null
```{r}
#null summary
na_summary(df,sort_by = "percent_complete") 
# we will just drop missing values because they weren't important 
df <- na.omit(df)
```

# 5. EDA
## 5.1 correlation
```{r}
nu_df <- df %>% select(where(is.numeric))
corrplot(cor(nu_df),method='pie')

cor_salary <- as.data.frame(cor(nu_df)[,'Salary'])
names(cor_salary) <- 'Salary'
cor_salary %>% arrange(desc(Salary))

```
* we want to see all the selected variables' correlations with the salary.
* we can see that PTS,FG,FGA are strongly correlated with salaries -> which might indicate that the more you score, the more money you will earn.
* it is quite surprising that 3P%, 2P%, TS% and FG% have little to do with salary 
* (we will verify this later in the machine learning part)

## 5.2 graphs
### 5.2.1 PTS VS Salary
**since PTS is most correlated with salary, let's look closely how the graph will look like**

```{r}
s_p <- ggplot(df,aes(Salary/1000000,PTS,color=Player)) + geom_point(alpha=0.5) + labs(x='Salary(million)', y='Total Points Scored', title='Salary vs Point')  +theme(plot.title = element_text(hjust=.5,size=14,face='bold'),legend.position = "none") + scale_y_continuous(breaks=seq(0,3000,500))+scale_x_continuous(breaks=seq(0,40,5)) 

ggplotly(s_p)
```
You can see that in general, pts and salary are positively correlated: if you scored more, you will get paid more.

### 5.2.2 Team Salary
***The National Basketball Association today announced that the Salary Cap has been set at $99.093 million for the 2017-18 season.***
```{r}
df %>% group_by(Team) %>% summarize(ts = sum(Salary)) %>% 
  ggplot(aes(reorder(Team,ts),ts/1000000)) + geom_col(color='black',fill=' tomato') + coord_flip()+labs(x='Total Salary(million)',y=NULL,title='Total Salary per Team') + theme(plot.title = element_text(size=14, hjust=.5)) + geom_hline(yintercept = 99.093,linetype='dotted',color='blue',size=1.5)
```
We can see that nearly 2/3 of NBA teams' salary are above the Salary Cap 99.093 million. It might be interesting to acquire team winning data and further compare them with total salary to see if they are correlated.

# 6. Machine Learning (Linear Regression)
## 6.1 Linear Regression
```{r}
sample <- sample.split(df$Salary,.7)

train <- subset(df, sample==T)
test <- subset(df, sample==F)

```
## 6.2 Training Model
```{r}
model <- lm(Salary~.,train[,c('Salary',"ORB","DRB","AST","STL","BLK","TOV","PTS")])
summary(model)
names(df)
```
***We only limited our predictors to ORB : PTS as these data are most readily available and we can acquire them online to predict a player's next season salary.***
* From the summary table, we can see that PTS,TOV,BLK,AST are four most important factors when it comes to NBA salary prediction.

## 6.3 Residuals
```{r}
res <- as.data.frame(residuals(model))
names(res) <- "residual"
ggplot(res,aes(residual)) + geom_histogram(fill='tomato',color='black')
```
**The residual histogram is normally distributed**

## 6.4 Test predictions
```{r}
Salary.predictions <- predict(model,test)
results <- cbind(Salary.predictions,test$Salary)
colnames(results) <- c('pred','actual')
results <- as.data.frame(results)
results

SSE = sum((results$pred - results$actual)^2)
SST = sum( (mean(df$Salary) - results$actual)^2)

R2 = 1 - SSE/SST
R2
```
***hm.. low R2 is not a good sign but let's keep exploring***

## 6.5 Predictions
***Let's predict Stephen Curry(the best player in the world:)) salary of next season**
```{r}
curry_stas <- data.frame(name = 'Stephen Curry', PTS = 1999,AST = 524,BLK=17,TOV=239 )

n_model <- lm(Salary~., df[,c('Salary','PTS','AST','BLK','TOV')])
outcome <- predict(n_model,curry_stas)
print(paste('pred:', outcome, 'real: 45780000'))
```











